---
title: "Data analysis"
author: "Group 6: Bianka R., Helene, Stine F.H., Nanna and Katrine"
date: "2/11/2020"
output: html_document
---

```{r setup, include=FALSE}
require(knitr)
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
library(tidyverse)
pacman::p_load(lme4, LmerTest)
ggplot2::theme_set(theme_bw())
library(DHARMa)

```

```{r load data}
#Samples <- read_csv("Exported_EyeLink_data/Samples_merged.csv") %>% 
 # mutate(GazeY = 1051-GazeY, Fix_MeanY = 1051-Fix_MeanY) %>% 
  #filter(Time<=41202)

Samples <- read_csv("Exported_EyeLink_data/Samples_merged.csv", col_types = cols(ParticipantID = col_character(), ParticipantGender = col_character(), EyeTracked = col_character(), Task = col_character(), SearchOrder = col_double(), ForagingType = col_character(), Trial = col_double(), Stimulus = col_character(), Video = col_character(), Time = col_double(), GazeX = col_double(), GazeY = col_double(), PupilSize = col_double(), FixationNo = col_double(), Fix_StartTime = col_double(), Fix_EndTime = col_double(), Fix_Duration = col_double(), Fix_MeanX = col_double(), Fix_MeanY = col_double(), Fix_MeanPupilSize = col_double(), SaccadeNo = col_double(), Sac_StartTime = col_double(), Sac_EndTime = col_double(), Sac_Duration = col_double(), Sac_StartX = col_double(), Sac_StartY = col_double(), Sac_EndX = col_double(), Sac_EndY = col_double(), Sac_PeakVelocity = col_double(), Sac_MeanVelocity = col_double(), Sac_Blink = col_logical(), Sac_Direction = col_character(), Sac_Amplitude = col_double()))
```

```{r what are the implications of modeling the data in a different way,  include=FALSE} 
## Subsetting the Forarging experiment
x = subset(Samples, Task == "Foraging", !is.na(SaccadeNo))

x <- x %>% group_by(ParticipantID, Stimulus, ForagingType, SaccadeNo) %>% summarise(Saccade_Amplitude = mean(Sac_Amplitude, na.rm = TRUE))

#We model
sac_ampl_gaus <- glmer(Saccade_Amplitude ~ 1 + ForagingType + 
        (1 + ForagingType | ParticipantID) +  
        (1 + ForagingType | Stimulus), data = x, family = gaussian(link=log))
summary(sac_ampl_gaus)
## The averages for count seems to be 0.9 but when doing a boxplot Riccardo saw that it's not 
# Because we put in the link = log the estimates are on a log scale. We need to invert these. Eksponentiate.  
exp(0.93+0.54)  #exp of intercept + slope. The average saccade length in the foraging???
exp(0.93) # exp of intercept. the average in the count task
exp(0.93+0.54) -  exp(0.93) #we subtract to get the difference
#count is the intercept foraging type. 
# The slope 0.5 needs to be inverted when we want to understand it

#The amplitude s in degrees where it's the angle it eyeball moves compared to the screen.

#We model
sac_ampl_id <- glmer(Saccade_Amplitude ~ 1 + ForagingType + 
        (1 + ForagingType | ParticipantID) +  
        (1 + ForagingType | Stimulus), data = x, family = gaussian(link=identity))
summary(sac_ampl_id)


p_1 <- predict(sac_ampl_gaus)
p_2 <- predict(sac_ampl_id)

#The distribution of the density based on the predictions of the gaussian lognomial model
plot(density(p_1))
#The distribution of the density of the gaussian model
plot(density(p_2)) #It does not have a long tale because it tries to normally distribute it. 
#plot(density(x$SaccadeNo))
summary(abs(p_1-x$Saccade_Amplitude))
summary(abs(p_2-x$Saccade_Amplitude))


dh_1 <- DHARMa::simulateResiduals(sac_ampl_gaus)


dh_2 <- DHARMa::simulateResiduals(sac_ampl_id)

plot(dh_1)
plot(dh_2)


```
With gaussian we are undermining the tales. 
With the lognomial it's not really better

Conclusions:
We tried to build a model that make conceptual sense. 
In which ways are the model failing us. 
We compared it with a gaussian model. Does it make sense to use a log model?
We find that there are longer saccade amplitude. This means a specific amount of degrees. 
We looked at the general distributions of the data. 0.5% R2m variance is explained. If we know the participants then 2.7%. 
The model can capture the general difference between conditions but it fails as capturing the single saccades. Some are exploratoty and some exploit. 

```{r Analysis for Assignment 1 Social Engagement}
#Social Engagement task
#We will look for differences in pupil size in engagement vs. observatonal only stimuli

## Subsetting the Social Engagement experiment
x = subset(Samples, Task == "SocialEngagement", !is.na(PupilSize))
#We take an average pupilsize for each fixation

#We make the video column into multiple columns
x <- separate(x, Video, c("Gender", NA, NA, "Orientation", "Ostensive"), remove = FALSE, sep = "_")

#We now group by participants and 
x <- x %>% group_by(Fix_MeanPupilSize, Trial, FixationNo, ParticipantID, Orientation, Ostensive, Gender) %>% 
  summarize(PupilSize = mean(PupilSize))

#We model
model <- glmer(PupilSize ~ 1 +  Orientation * Ostensive +
        (1 + Orientation * Ostensive | ParticipantID), data = x, family = gaussian(link=log))
summary(model) #We use log because we believe it will reflect the data better. 

#Here we get no significant interaction so we model again without
model <- glmer(PupilSize ~ 0 +  Orientation + Ostensive +
        (0 + Orientation + Ostensive | ParticipantID), data = x, family = gaussian(link=log))
summary(model)


x$Ostensive <- as.factor(x$Ostensive)
x$Ostensive <- relevel(x$Ostensive, ref = "+o")

model <- glmer(PupilSize ~ 0 +  Orientation + Ostensive +
        (0 + Orientation + Ostensive | ParticipantID), data = x, family = gaussian(link=log))
summary(model)


#We get a significant effect of the diverted orientation. When doing it with link = identity we do not get anything significant. 

#We need to plot it
ggplot(x, aes(FixationNo, PupilSize, color = Orientation)) + geom_smooth(method = "lm")

#We plot with participant ID as well
ggplot(x, aes(FixationNo, PupilSize, color = ParticipantID, fill = Orientation)) + geom_smooth(method = "lm", se = F, aes(linetype=Orientation))
```


```{r Assignment 1 Foraging experiment}
## Subsetting the Forarging experiment
x = subset(Samples, Task == "Foraging", !is.na(SaccadeNo))

x <- x %>% group_by(ParticipantID, Stimulus, ForagingType, SaccadeNo) %>% summarise(Saccade_Amplitude = mean(Sac_Amplitude, na.rm = TRUE))

#We model
sac_ampl_gaus <- glmer(Saccade_Amplitude ~ 1 + ForagingType + 
        (1 + ForagingType | ParticipantID) +  
        (1 + ForagingType | Stimulus), data = x, family = gaussian(link=log))
summary(sac_ampl_gaus)
# We plot
ggplot(x, aes(SaccadeNo, Saccade_Amplitude, color=ParticipantID))+
  geom_smooth(aes(linetype=ForagingType), method="lm",se=F) + facet_wrap(.~Stimulus)

ggplot(x, aes(SaccadeNo, Saccade_Amplitude, color=ParticipantID))+
  geom_smooth(aes(linetype=ForagingType), method="lm",se=F)
```



```{r Estimating events manually,  include=FALSE}
#We use only x,y and time

Samples2 <- Samples %>% 
  mutate(Velocity = abs((lead(GazeX)-GazeX)/Time*1000)) #We multiply with 1000 to go into seconds? We get very big changes so we will smooth to make it more realistic
#We use pythagoras a^2 + b^2 = c^2. The lead gaze x minus the previous
#We use abs because we don't want the negative values. 
#We can't trust these numbers because they are pixels per second and not degrees. 

#The values are quite low when it's durring a fixation. 

#Now we make an algorithm
Samples2 <- Samples %>% 
  mutate(Velocity = abs((lead(GazeX)-GazeX)/Time*1000)) %>%
  mutate(Event = ifelse(Velocity >= 1.5, "Saccade", "Fixation"))

#We need some kind of smoothing. If mostly saccades with a few fixations inbetween we want to call the whole chunk a saccade
Samples2 <- Samples %>% 
  mutate(Velocity = abs((lead(GazeX)-GazeX)/Time*1000)) %>%
  mutate(Event = ifelse(Velocity >= 1.5, "Saccade", "Fixation")) %>% 
  filter(Velocity != "Inf") %>%
  mutate(Velocity_smooth = (cumsum(Velocity) - lag(cumsum(Velocity), n = 20))/20) %>%
  mutate(Event_smooth = ifelse(Velocity_smooth >= 0.5, "Saccade", "Fixation"))


ggplot(data = Samples2[Samples2$ParticipantID == "F2" & 
  Samples2$Trial == 1,], 
  aes(Time, GazeX, color = ParticipantID)) +
  geom_point()


```

